Estimation of the road traffic sound levels in urban areas based on non-negative matrix factorization techniques

### 2018-09-06 revision/gloaguenNmfTraffic_v2.tex
an-bs and intro

In this paper, several techniques based on the non-negative matrix  factorization framework are studied in this application scenario, on a corpus of simulated sound scenes. The task is to the best of our knowledge never been considered in the literature. We propose an experimental protocol to validate the studied approaches that complies with standard reproducible research recommendations. The results show the interest of our proposed approach (a NMF based on the adjustment of a learned dictionary directly on the urban sound mixtures) for such sound environments as it improves the estimation of the road traffic sound level compared to basic methods such as a low-pass filter or supervised NMF.

@@

With the introduction of the European Directive 2002/\-EC/49, cities over 100 000 inhabitants have to produce road traffic noise maps. These maps depict the sound level distribution over the city and an estimation of the number of city dwellers exposed to high noise levels. These maps play both an important communication role and help drawing up action plans to reduce noise exposure. Road traffic noise maps are the result of a simulation process based on the estimation of the traffic density on the main roads and the use of sound propagation modelling. They express as output $L_ {DEN}$ and $L_N$ values, which are \textit{Day-Evening-Night} and \textit{Night} equivalent A-weighted sound levels respectively. Although very useful, the produced noise maps introduce lot of uncertainty generated by the numerical tools \cite{van_leeuwen_noise_2015} or by the different calculation methodologies used \cite{leroy_uncertainty_2010}\cite{garg_critical_2014}, despite the long data collection and calculation times. In addition, the usual road traffic noise maps are static, aggregating the exposure into indicators $L_{DEN}$ and $L_N$, that ignore the sound levels evolution throughout the day.
The use of acoustic measurements could facilitate their updating or even the generation of dynamic maps \cite{wei_dynamic_2016}. These measurements can be performed at fixed stations spread all over the cities \cite{Mioduszewski} \cite{mietlicki2012innovative}, which would make the long-term evolution of the traffic noise available. It can also be performed with  mobile stations \cite{can_exploring_2012} \cite{manvell2004sadmam} covering a larger area with fewer sensors but also sparse time periods.

@@

Prior to data assimilation, the issue of the correct estimation of the traffic sound level from acoustic measurements begin to be studied \cite{leiba2017large,socoro_anomalous_2017}. As the urban sound environment is a complex environment gathering lots of different sounds (car passages, voices, whistling bird, car horn, airplanes\dots) that overlap, the traffic sound level estimation based on measurements is not a trivial task.

@@

Recent work is mostly focused on the detection or recognition tasks of environmental sounds \cite{heittola_sound_2011}, \cite{defreville_automatic_2006}, \cite{dufaux_automatic_2000}, \cite{chu_environmental_2009}. A two-step process is generally followed: describe the audio files with a set of features (Spectral Centroid, harmonicity, Mel-Frequency Cepstral Coefficient \dots) and classify them with the help of classifiers (Support Vector Machines, Gaussian Mixture Models, Hidden Markov Model, Artifical Neural Networks). A description of these features and classifiers can be found in \cite{cowling_comparison_2003} and their applications can be found in \cite{shen_environmental_2012}, \cite{beritelli_pattern_2008}, \cite{couvreur_automatic_2004}.

The main issue in the detection or recognition tasks is the overlap of environmental sounds. Although near major roads, traffic is predominant, there are many places where it overlaps with other sound sources which contribute significantly to the overall sound levels. To circumvent this issue, Socor\'o et al. propose to suppress time frames where there is significant overlap by considering an Anomalous Noise Events Detector \cite{socoro_anomalous_2017}. It consists in detecting the unwanted sound sources from labeled recordings, \textit{i.e.} that are not related to the traffic component. Those time frames are then discarded in order not to take them into account during the estimation of the traffic sound level.
An alternative approach that we will follow in this paper is to consider the blind source separation paradigm to reliably estimate the traffic noise level, see Figure \ref{fig:diagram}. It consists in separating the contribution of the traffic from the other sources within a polyphonic scene. One major advantage of following such approach is that the estimate is continuously available, making the approach applicable in a wide range of urban areas, even where the traffic noise is relatively low compared to the remaining contributions.\\

In an urban environment context, source separation can be achieved with the help of acoustic microphone arrays and beamforming \cite{saruwatari2003blind}. However, this approach requires spreading multiple microphones arrays in cities that is very expensive (even with low cost microphones) and time-consuming for calibration and maintenance. This method is then not considered here to be deployed all over cities. On the contrary, monophonic sensor networks need less microphones but the main challenge is to succeed to estimate correctly the road traffic from only one signal in which all kind of sound sources can be present. A convenient method for this is the  Non-negative Matrix Factorization (NMF) technique \cite{lee_learning_1999}. When considering audio as input, it usually consists in approximating the magnitude spectrogram of an audio file by the product of two low rank matrices, one representing the components of interest and the other the contribution at a given time of those components to approximate the input magnitude spectrogram \cite{smaragdis_non-negative_2003} \cite{wilson_speech_2008}. In the audio processing domain, NMF has already been employed for the source separation task of monaural signals of speech and music \cite{wang_musical_2005} \cite{wilson_speech_2008}. By design, this method deals reasonably well with the overlapping sound sources as soon as the overlap can be resolved on the time/frequency plane.

Closer to our application scenario, NMF has been considered in \cite{dikmen2013sound} where coupled NMF is used as a sound event detector. The dictionary learning is based on the spectrogram of audio recordings and on their annotations which share the same activation matrix. If the performance is satisfying (especially for non-negative SNRs), this approach has the advantage of allowing the use of audio recordings consisting of multiple and overlapping sound sources and reducing the ambiguity of unsupervised learning. However, if the sound event can be well detected, no information on the quality of the detected signal is mentioned.
Similarly, in \cite{mesaros_sound_2015}, NMF is used in the same task but with real life recordings. The authors raised the question of the dictionary size reduction after the learning phase: cluster the elements and keep the full spectrum or use mel spectrum representation and keep the full learnt dictionary? They conclude on the efficiency of each approach and suggere to combine both to deal with large databases. Also, Innami and Kasai \cite{satoshi_innami_nmf-based_2012} proposed to perform NMF on simulated audio files as a source separation method. Their tool is composed of two steps by 1) separating the sound background from the events and 2) by isolating the events using spectral features using a $k$-means procedure.
On a preliminary study \cite{gloaguen2016estimating}, supervised Non-Negative Matrix Factorization has been considered, on simple simulated sound scenes, as a source separation method to extract the road traffic component (which include continuous traffic noise and passing car) in order to estimate its sound level.

Here, we extend this approach with more flavors of NMF on more complex sound scenes. This corpus of simulated sound scenes is created from a built-up sound database composed of a high number of diverse sound samples in order to encompass a wide variety of sound environment with a variable presence of road traffic noise. The use of simulated sound scenes allows rigorous experimental validation as it offers a high level of control on the design of the scenes and the knowledge of the exact contribution of the traffic component ($L_{p,traffic}$), which would be difficult to extract from a recording of an urban scene.
The aim is to find the form of NMF that gives the minimal error on the reconstruction of the traffic noise signal on the whole corpus to be applied to all the different possible urban noise environments.
We demonstrate that supervised NMF and semi supervised NMF approaches have some interests but fail to give satisfactory results for the application at hand. We thus introduce another scheme called thresholded initialized NMF that makes good use of prior knowledge about the source of interest, in our case the traffic noise, but also generalizes well to several kinds of urban areas and to traffic to interference ratio ($TIR$).

>>>>

In this paper, several techniques based on the non-negative matrix  factorization framework are studied in this application scenario, on a corpus of simulated sound scenes. The task is to the best of our knowledge never been considered in the literature. We propose an experimental protocol to validate the studied approaches that complies with standard reproducible research recommendations. The results show the interest of our proposed approach %(a NMF based on the adjustment of a learned dictionary directly on the urban sound mixtures)
for such sound environments.% as it improves the estimation of the road traffic sound level compared to basic methods such as a low-pass filter or supervised NMF.

@@

With the introduction of the European Directive 2002/\-EC/49, cities over 100 000 inhabitants have to produce road traffic noise maps. These maps depict the sound level distribution over the city and an estimation of the number of city dwellers exposed to high noise levels. These maps play both an important communication role and help drawing up action plans to reduce noise exposure. Road traffic noise maps are the result of a simulation process based on the estimation of the traffic density on the main roads and the use of sound propagation modelling. They express as output $L_ {DEN}$ and $L_N$ values, which are \textit{Day-Evening-Night} and \textit{Night} equivalent A-weighted sound levels respectively. Although very useful, the produced noise maps exhibit a large amount of uncertainty generated by the numerical tools \cite{van_leeuwen_noise_2015} or by the different calculation methodologies used \cite{leroy_uncertainty_2010}\cite{garg_critical_2014}. It also requires a labor intensive data collection and calculation times. The result is usually that road traffic noise maps are static, aggregating the exposure into indicators $L_{DEN}$ and $L_N$, that ignore the sound levels evolution throughout the day.
The use of acoustic measurements could facilitate their updating or even the generation of dynamic maps \cite{wei_dynamic_2016}. These measurements can be performed at fixed stations spread all over the cities \cite{Mioduszewski} \cite{mietlicki2012innovative}, which would make the long-term evolution of the traffic noise available. It can also be performed with  mobile stations \cite{can_exploring_2012} \cite{manvell2004sadmam} covering a larger area with fewer sensors but also sparse time periods.

@@

Prior to data assimilation, the issue is to correctly estimate the traffic sound level from acoustic measurements as studied in \cite{leiba2017large,socoro_anomalous_2017}. As the urban sound environment is a complex environment gathering lots of different sounds (car passages, voices, whistling bird, car horn, airplanes\dots) that overlap, the estimation of the traffic sound level  based on acoustic measurements is not a trivial task.

@@

Recent work is mostly focused on the detection or recognition tasks of environmental sounds \cite{heittola_sound_2011}, \cite{defreville_automatic_2006}, \cite{dufaux_automatic_2000}, \cite{chu_environmental_2009}. A two-step process is generally followed: describe the audio files with a set of features (Spectral Centroid, harmonicity, Mel-Frequency Cepstral Coefficient \dots) and classify them with the help of classifiers (Support Vector Machines, Gaussian Mixture Models, Hidden Markov Model, Neural Networks) \cite{cowling_comparison_2003} and can be applied to a variety of tasks \cite{shen_environmental_2012}, \cite{beritelli_pattern_2008}, \cite{couvreur_automatic_2004}.

The main issue with the detection or recognition tasks is the overlap of environmental sounds. Although near major roads, traffic is predominant, there are many places where it overlaps with other sound sources which contribute significantly to the overall sound levels. To circumvent this issue, Socor\'o et al. propose to suppress time frames where there is significant overlap by considering an Anomalous Noise Events Detector \cite{socoro_anomalous_2017}. It consists in detecting the unwanted sound sources from labeled recordings, \textit{i.e.} that are not related to the traffic component. Those time frames are then discarded in order not to take them into account during the estimation of the traffic sound level.
An alternative approach that we will follow in this paper is to consider the blind source separation paradigm to reliably estimate the traffic noise level, see Figure \ref{fig:diagram}. It consists in separating the contribution of the traffic from the other sources within a polyphonic scene. One major advantage of following such approach is that the estimate is  available continuously, making the approach applicable in a wide range of urban areas, even where the traffic noise is relatively low compared to the remaining contributions.\\

In an urban environment context, source separation can be achieved with the help of acoustic microphone arrays and beamforming \cite{saruwatari2003blind}. However, this approach requires the spreading of multiple microphones arrays in cities that is very expensive (even with low cost microphones) and time-consuming for calibration and maintenance.  Monophonic sensor networks need less microphones but the main challenge is to succeed to estimate correctly the level of the road traffic from a sound mixture in which all kind of sound sources can be present. A convenient method for this is the  Non-negative Matrix Factorization (NMF) technique \cite{lee_learning_1999}. When considering audio as input, it usually consists in approximating the magnitude spectrogram of an audio file by the product of two low rank matrices, one representing the components of interest and the other the contribution at a given time of those components to approximate the input magnitude spectrogram \cite{smaragdis_non-negative_2003} \cite{wilson_speech_2008}. In the audio processing domain, NMF has already been employed for the source separation task of monaural signals of speech and music \cite{wang_musical_2005} \cite{wilson_speech_2008}. By design, this method deals reasonably well with the overlapping sound sources as soon as the overlap can be mostly resolved on the time/frequency plane.

Closer to our application scenario, NMF has been considered in \cite{dikmen2013sound} where coupled NMF is used as a sound event detector. The dictionary learning is based on the spectrogram of audio recordings and on their annotations which share the same activation matrix. If the performance is satisfying (especially for non-negative SNRs), this approach has the advantage of allowing the use of audio recordings consisting of multiple and overlapping sound sources and reducing the ambiguity of unsupervised learning. However, if the sound event can be well detected, no information on the quality of the detected signal is mentioned.
Similarly, in \cite{mesaros_sound_2015}, NMF is used in the same task but with real life recordings. The authors raised the question of the dictionary size reduction after the learning phase: cluster the elements and keep the full spectrum or use mel spectrum representation and keep the full learnt dictionary? They conclude on the efficiency of each approach and propose to combine both to deal with large databases. Also, Innami and Kasai \cite{satoshi_innami_nmf-based_2012} proposed to perform NMF on simulated audio files as a source separation method. Their tool is composed of two steps by 1) separating the sound background from the events and 2) by isolating the events using spectral features using a $k$-means procedure.
On a preliminary study \cite{gloaguen2016estimating}, supervised Non-Negative Matrix Factorization has been considered, on simple simulated sound scenes, as a source separation method to extract the road traffic component (which include continuous traffic noise and passing car) in order to estimate its sound level.

Here, we extend this approach with more flavors of NMF on more challenging sound scenes. The corpus of simulated sound scenes considered here is created from a built-up sound database composed of a large number of diverse sound samples in order to encompass a wide variety of sound environment with a variable presence of road traffic noise. The use of simulated sound scenes allows rigorous experimental validation as it offers a high level of control on the design of the scenes and the knowledge of the exact contribution of the traffic component ($L_{p,traffic}$), which would be difficult to extract from a recording of an urban scene.
The aim is to find the type of NMF that gives the minimal error on the estimation of the traffic noise signal level on the whole corpus.% to be applied to all the different possible urban noise environments.
We demonstrate that supervised NMF and semi supervised NMF approaches have some interests but do not give satisfactory results for the application at hand. We thus introduce another scheme called thresholded initialized NMF that makes good use of prior knowledge about the source of interest, in our case the traffic noise, but also generalizes well to several kinds of urban areas and to traffic to interference ratio ($TIR$).
